---
title: "dwc_mapping_fixed_terms"
subtitle: "Mica -  Agouti Muskrat and Coypu Camera trap Occurrences collected in Flanders, the Netherlands and Germany"
author: "Dimitri Brosens"
date: "12-1-2021"
output: html_document  
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
```

# Read source interim data

This script ass a number of (often used) fixed terms to an already existing .csv file.For this script is, the original file is dwc_occurrence and dwc_multimedia from 'Agouti Muskrat occurrences'.
As the data is not RAW it we will start with the interim data. For reading the file we use the most appropriate function

### input_interim <- read.csv2(here::here("datasets", "mica-rato-occurrences", "data", "raw", "RATOgegevensbackoffice2020V2.txt" ), sep = ";")

# For excel files

### input_interim <- read_excel(here::here("datasets", "mica-rato-occurrences", #"data", "raw", "datasetNAme.xlsx"))

# For delimited files, who needs more specification

###input_interim <- read_delim(file = here::here("datasets", "mica-agouti-occurrences", "data", "interim", "dwc_occurrence.csv"), ",", escape_double = FALSE, trim_ws = TRUE )


```{r}
#for csv files
input_interim <- read_csv(here::here("datasets", "mica-agouti-occurrences", "data", "interim", "dwc_occurence.csv" ))



```

Preview data:

```{r}
input_interim %>% head(n = 5)
```


# Occurrence core / event core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim  #for occurrence core datasest
# event <- input_interim       #for event      core datasets
```


# Fixed Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml). 

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records). For the term order check [tdwg](https://dwc.tdwg.org/terms/)

# Record level

### type
here already in the source data

```{r}
occurrence %<>% mutate(dwc_type = type)
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 

```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```


### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```



### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Mica -  Agouti Muskrat and Coypu Camera trap Occurrences collected in Flanders (Belgium), the Netherlands and Germany") 
```
# occurrence

### basisOfRecord

not always fixed term

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = basisOfRecord)
```
### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = occurrenceID)
```

### recordedBy

not always fixed term

```{r}
occurrence %<>% mutate(dwc_recordedBy = 'Agouti')
```
### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = individualCount)
```

### sex

```{r}
occurrence %<>% mutate(dwc_sex = sex)
```


### lifestage

```{r}
occurrence %<>% mutate(dwc_lifeStage = lifeStage)
```


### behavior

```{r}
#occurrence %<>% mutate(dwc_behavior = behavior)
```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = occurrenceStatus)
```

# Event

### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = eventID)
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = eventDate)
```


## samplingProtocol

not always fixed term

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = samplingProtocol) 

```


# Location

### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = locationID)
```


### continent


```{r}
occurrence %<>% mutate(dwc_continent = "Europe") 
```
### locality

```{r}
occurrence %<>% mutate(dwc_locality = locality)
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = decimalLatitude)
```


###decimalLongitude


```{r}
occurrence %<>% mutate(dwc_decimalLongitude = decimalLongitude)
```


### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = geodeticDatum) # e.g. "Belgium = BE"
```


### coordinateUncertaintyInMeters



```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "200") 
```

# identification

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = identifiedBy)
```


### identificationRemarks


```{r}
occurrence %<>% mutate(dwc_identificationRemarks = identificationRemarks)
```


# Taxon

The following terms contain information about the taxon:


### scientificName


```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = kingdom)  #
```
### vernacularName


```{r}
occurrence %<>% mutate(dwc_vernacularName = vernacularName)
```


### taxonRank

not always fixed term

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### nomenclaturalCode



```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```


## Post-processing

```{r}

occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
``` 


```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-agouti-occurrences", "data", "processed", " occurrence.csv"), na = "")
```

# data quality tests

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the dataset:

```{r}
parsed_names <- occurrence %>%
  distinct(scientificName) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Preview data:

```{r}
parsed_names %>% head(n = 62)
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

```{r}
parsed_names <- input_interim %>%
  distinct(scientificName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
head(parsed_names, n = 30)
write.csv(parsed_names, "../data/processed/speciesList.csv", append = FALSE, na = "",  row.names = FALSE, fileEncoding = "UTF-8")
```

```{r}
#end
```

