---
title: "dwc_mapping_fixed_terms"
subtitle: "Mica -  Agouti Muskrat and Coypu Camera trap Occurrences collected in Flanders, the Netherlands and Germany"
author: "Dimitri Brosens"
date: "12-1-2021"
output: html_document  
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
```

# Read source interim data

This script ass a number of (often used) fixed terms to an already existing .csv file.For this script is, the original file is dwc_occurrence and dwc_multimedia from 'Agouti Muskrat occurrences'.
As the data is not RAW it we will start with the interim data. For reading the file we use the most appropriate function

### input_interim <- read.csv2(here::here("datasets", "mica-rato-occurrences", "data", "raw", "RATOgegevensbackoffice2020V2.txt" ), sep = ";")

# For excel files

### input_interim <- read_excel(here::here("datasets", "mica-rato-occurrences", #"data", "raw", "datasetNAme.xlsx"))

# For delimited files, who needs more specification

###input_interim <- read_delim(file = here::here("datasets", "mica-agouti-occurrences", "data", "interim", "dwc_occurrence.csv"), ",", escape_double = FALSE, trim_ws = TRUE )


```{r}
#for csv files
input_interim <- read_csv(here::here("datasets", "mica-agouti-occurrences", "data", "interim", "dwc_occurence.csv" ))



```

Preview data:

```{r}
input_interim %>% head(n = 5)
```


# Occurrence core / event core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim  #for occurrence core datasest
# event <- input_interim       #for event      core datasets
```


# Fixed Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml). 

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records). For the term order check [tdwg](https://dwc.tdwg.org/terms/)

# Record level

### type
here already in the source data

```{r}
#occurrence %<>% mutate(type = "Event")
```

### language

```{r}
occurrence %<>% mutate(language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/") 

```

### rightsHolder

```{r}
occurrence %<>% mutate(rightsHolder = "INBO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### collectionID

```{r}
###occurrence %<>% mutate(collectionID = "insertCollectionID") 
```

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(institutionCode = "INBO") # e.g. "INBO"
```

### collectionCode
not applicable for non collection datasets

```{r}
# occurrence %<>% mutate(collectionCode = "insertCollectionCode") 
```


### datasetName

```{r}
occurrence %<>% mutate(datasetName = "Mica -  Agouti Muskrat and Coypu Camera trap Occurrences collected in Flanders (Belgium), the Netherlands and Germany") 
```

### basisOfRecord

not always fixed term

```{r}
# occurrence %<>% mutate(basisOfRecord = "insert basisOfREcord")
```

### informationWithHeld

not always fixed term

```{r}
# occurrence %<>% mutate(basisOfRecord = "insert informationwithheld")
```


### dataGeneralizations

notalways fixed term

```{r}
# occurrence %<>% mutate(basisOfRecord = "insert data generalizations")
```
# Occurrence

### recordedBy

not always fixed term

```{r}
occurrence %<>% mutate(recordedBy = 'Agouti')
```

### organismQuantityType

not always fixed term

```{r}
occurrence %<>% mutate(organismQuantityType = 'Agouti')
```

# Event

### habitat

not always fixed term

```{r}
#occurrence %<>% mutate(habitat = 'insert habitat')
```

## samplingProtocol

not always fixed term

```{r}
#occurrence %<>% rename(samplingProtocol = 'insert samplingProtocol') %>%

```


### samplesizeUnit

not always fixed term

```{r}
#occurrence %<>% rename(samplingEffort = 'insert sampleSizeUnit')

```

### samplingEffort

not always fixed term

```{r}
#occurrence %<>% rename(samplingEffort = 'insert samplingEffort')

```


# Location

### continent

not always fixed term

```{r}
occurrence %<>% mutate(continent = "Europe") 
```

### country

not used

### countryCode

```{r}
occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```


### geodeticDatum

```{r}
#occurrence %<>% mutate(geodeticDatum = "WGS84") # e.g. "Belgium = BE"
```


### coordinateUncertaintyInMeters

not always a fixed term

```{r}
occurrence %<>% mutate(coordinateUncertaintyInMeters = "insertCoordinateUncertaintyInMeters") 
```

### coordinatePrecission

Not always fixed term
Not used

### verbatimCoordinateSystem

```{r}
#occurrence %<>% mutate(verbatimCoordinateSystem = "insertVerbatimCoordinateSystem") # e.g. "lambert coordinates"
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "insertVerbatimSRS") # e.g. "Belgium Datum 1972"
```


### georeferenceRemarks

not always fixed term

```{r}
#occurrence %<>% mutate(georeferenceRemarks = "insert georeferenceRemarks") # e.g. "Belgium = BE"
```

# Taxon

The following terms contain information about the taxon:



### identifiedBy

not always fixed term

```{r}
#occurrence %<>% mutate(identifiedBy = 'insert identifiedBy ')  #e.g. "RATO"
```


### kingdom

```{r}
occurrence %<>% mutate(kingdom = "Animalia")  #
```


### taxonRank

not always fixed term

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

### nomenclaturalCode

not alwasy fixed term

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```


## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
occurrence %<>% select(-c(eventRemarks)) 
          # %<>% #remove collection columns
          #  rename(scientificName = verbatimScientificName) 
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns

```{r}
col_order <- c( "type","language", "license","rightsHolder","accessRights","datasetID"
               ,"institutionCode", "datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount", "sex" ,"lifeStage", "behavior","occurrenceStatus","occurrenceRemarks"
               ,"eventID", "eventDate", "samplingProtocol"
               ,"locationID", "locality" ,"continent","countryCode" ,"decimalLatitude","decimalLongitude","geodeticDatum", "coordinateUncertaintyInMeters"
	             ,"identifiedBy" ,"dateIdentified", "identificationRemarks"
	             ,"scientificName", "kingdom", "taxonRank", "vernacularName" ,"nomenclaturalCode"
               )

occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-rato-occurrences", "data", "processed", " occurrence.csv"), na = "")
```

# data quality tests

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the dataset:

```{r}
parsed_names <- occurrence %>%
  distinct(scientificName) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Preview data:

```{r}
parsed_names %>% head(n = 62)
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

```{r}
parsed_names <- input_interim %>%
  distinct(scientificName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
head(parsed_names, n = 30)
write.csv(parsed_names, "../data/processed/speciesList.csv", append = FALSE, na = "",  row.names = FALSE, fileEncoding = "UTF-8")
```

```{r}
#end
```

