---
title: "Darwin Core Mapping"
subtitle: "MICA - Muskrat and Coypu camera trap occurrences collected in Belgium, the Netherlands and Germany"
author: "Dimitri Brosens, Peter Desmet"
date: "`r Sys.Date()`"
output: html_document  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(datapackage)    # To read the source data
```

# Quick timezone conversion

Most of the Darwin Core mapping is done by running queries in sqlite. But timezones get ignored when reading data into sqlite. R converts them nicely to UTC. So the following code reads, converts and writes the raw data files, which can then be uploaded in sqlite:

```{r eval = FALSE}
dp <- read_package("https://zenodo.org/record/4893244/files/datapackage.json")
deployments <- read_resource(dp, "deployments")
multimedia <- read_resource(dp, "multimedia")
observations <- read_resource(dp, "observations")
```

```{r eval = FALSE}
write_csv(deployments, here::here("datasets", "mica-agouti-occurrences", "data", "raw", "deployments.csv"), na = "")
write_csv(multimedia, here::here("datasets", "mica-agouti-occurrences", "data", "raw", "multimedia.csv"), na = "")
write_csv(observations, here::here("datasets", "mica-agouti-occurrences", "data", "raw", "observations.csv"), na = "")
```

# Read source data

Data are exported from Agouti in the [Camtrap DP](https://gitlab.com/oscf/camtrap-dp) format. The main transformation to Darwin Core is done using the generic camtrap-dp to dwc guidelines in [dwc-for-biologging](https://github.com/tdwg/dwc-for-biologging/tree/master/derived/camtrap-dp). **We thus start from Darwin Core formatted files** (in `data/interim`), that we enhance mostly with static terms that are not yet included.

```{r}
occurrence <- read_csv(here::here("datasets", "mica-agouti-occurrences", "data", "interim", "dwc_occurrence.csv"))
multimedia <- read_csv(here::here("datasets", "mica-agouti-occurrences", "data", "interim", "dwc_multimedia.csv"))
```

Preview data:

```{r}
occurrence %>% head()
```

# Darwin Core mapping

## Occurrence core

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = type)
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # NEW TERM
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") # NEW TERM
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # NEW TERM
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") # NEW TERM
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/5tb6ze") # NEW TERM
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # NEW TERM
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "MICA - Muskrat and coypu camera trap observations in Belgium, the Netherlands and Germany") # NEW TERM
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = basisOfRecord)
```

### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = occurrenceID)
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = individualCount)
```

### sex

```{r}
occurrence %<>% mutate(dwc_sex = sex)
```

### lifestage

```{r}
occurrence %<>% mutate(dwc_lifeStage = lifeStage)
```

### behavior

```{r}
# EXCLUDED: always empty
```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = occurrenceStatus)
```

### organismID

```{r}
# EXLUCED: always empty
```

### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = eventID)
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = eventDate)
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = samplingProtocol)
```

### eventRemarks

```{r}
# EXLUCED: not relevant
```

### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = locationID)
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe") # NEW TERM
```

### locality

```{r}
occurrence %<>% mutate(dwc_locality = locality)
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.3f",
  round(decimalLatitude, digits = 3)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.3f",
  round(decimalLongitude, digits = 3)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = geodeticDatum)
```

### coordinateUncertaintyInMeters

Coordinates are rounded to 3 decimals = 111m maximum, see <https://github.com/inbo/mica-occurrences/issues/6>.

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "111") # NEW TERM
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = identifiedBy)
```

### dateIdentified

```{r}
occurrence %<>% mutate(dwc_dateIdentified = dateIdentified)
```

### identificationRemarks

```{r}
occurrence %<>% mutate(dwc_identificationRemarks = identificationRemarks)
```

### identificationVerificationStatus

```{r}
occurrence %<>% mutate(dwc_identificationVerificationStatus = identificationVerificationStatus)
```

### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = recode(scientificName,
  "Canis lupus" = "Canis lupus familiaris" # There are no wolves in dataset
))
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = kingdom)
```

### vernacularName

```{r}
occurrence %<>% mutate(dwc_vernacularName = recode(vernacularName,
  "gray wolf" = "dog" # There are no wolves in dataset
))
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = case_when(
  str_count(dwc_scientificName, " ") == 2 ~ "subspecies", # Including dogs
  str_count(dwc_scientificName, " ") == 1 ~ "species",
  dwc_scientificName %in% c("Anas", "Anser") ~ "genus",
  dwc_scientificName %in% c("Columbidae", "Soricidae") ~ "family",
  dwc_scientificName %in% c("Rodentia") ~ "order"
)) # NEW TERM
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # NEW TERM
```

## Multimedia extension

Map the data to [Simple Multimedia extension](https://rs.gbif.org/extension/gbif/1.0/multimedia.xml).

### occurrenceID

```{r}
multimedia %<>% mutate(dwc_occurrenceID = occurrenceID)
```

### type

```{r}
multimedia %<>% mutate(dwc_type = type)
```

### format

```{r}
multimedia %<>% mutate(dwc_format = format)
```

### identifier

```{r}
multimedia %<>% mutate(dwc_identifier = identifier)
```

### created

```{r}
multimedia %<>% mutate(dwc_created = created)
```

### license

```{r}
multimedia %<>% mutate(dwc_license = "https://creativecommons.org/licenses/by/4.0/") # NEW TERM
```

### rightsHolder

```{r}
multimedia %<>% mutate(dwc_rightsHolder = "INBO") # NEW TERM
```

# Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
multimedia %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
colnames(multimedia) <- str_remove(colnames(multimedia), "dwc_")
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-agouti-occurrences", "data", "processed", "occurrence.csv"), na = "")
write_csv(multimedia, here::here("datasets", "mica-agouti-occurrences", "data", "processed", "multimedia.csv"), na = "")
```
