---
title: "Darwin Core Mapping"
subtitle: "MICA - Muskrat occurrences collected by VMM in Flanders, Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(rgbif)          # To use GBIF services
library(uuid)           # To generate UUID's
```

# Read source data

```{r}
input_data <- read_excel(here::here("datasets", "mica-vmm-occurrences", "data", "raw", "VangstenT08_34_50.xlsx"))
```

Preview data:

```{r}
input_data %>% head()
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% 
  remove_empty("rows") %<>%
  clean_names()
```

# Retrieve scientific names based on the vernacular name

The raw data does not contain scientific names. We use the GBOF backbone taxonomy to retrieve scientific names based on vernacular names. Manual actions are still needed as the returned information might not be 100% correct.

See original vernacular names:

```{r}
input_data %>%
  distinct(sporen_waarnemingen_naam)
```

Create column `vernacularName`, remove data that is not specific enough and clean some names:

```{r}
input_data <-
  input_data %>%
  mutate(vernacularName = sporen_waarnemingen_naam) %>%
  filter(vernacularName != 'Vis') %>%
  filter(vernacularName != 'Andere') %>%
  filter(vernacularName != 'Eend') %>%
  filter(vernacularName != 'Marterachtige') %>%
  mutate(vernacularName = recode(vernacularName,
    "Muskusr. > 400gr" = "Muskusrat",
    "Doodaars" = "Dodaars"
  ))
```

Use GBIF function to find scientific names based on cleaned vernacular names:

```{r}
get_vernacular_name <- function(vn) {
 names <- name_lookup(
   vn,
   datasetKey = "d7dddbf4-2cf0-4f39-9b2a-bb099caae36c",
   limit = 1
 )$data # this returns the most likely taxon
 if (nrow(names) > 0) {
   names$scientificName
 } else {
   NA_character_
 }
}

scientific_names <- input_data %>%
  distinct(vernacularName) %>%
  mutate(scientificName = map_chr(vernacularName, get_vernacular_name))
```

Join scientific names with original dataframe:

```{r}
input_data <-
  input_data %>%
  right_join(scientific_names, by = "vernacularName")

input_data %>% distinct(vernacularName, scientificName)
```

Correct wrong scientific names:

```{r}
input_data %<>% mutate(scientificName = recode(scientificName,
  "Zapornia pusilla intermedia (Hermann, 1804)" = "Gallinula chloropus (Linnaeus, 1758)"
))
```

# Darwin Core mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
occurrence <- input_data
```

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "VMM")
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "insert doi")
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "VMM")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "MICA - Muskrat occurrences collected by VMM in Flanders, Belgium")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### occurrenceID

Since `registratie_id` is not unique for all records, we create an `occurrenceID` based on row number:

```{r}
occurrence <-
  occurrence %>%
  rowid_to_column(var = "row_id") %>%
  mutate(dwc_occurrenceID = str_c("MICA:VMM:OCC:", row_id))
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = str_c("VMM team ", team_naam))
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = vangst_aantal) 

```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = case_when(
  dwc_individualCount > 0 ~ "present",
  dwc_individualCount == 0 ~ "absent"
))
```

### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = str_c("MICA:VMM:EV:", registratie_id))
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = as.character(dag))
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = "rat trap") 
```

### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = locatie_id)
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe")
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE")
```

### stateProvince

```{r}
occurrence %<>% mutate(dwc_stateProvince = recode(provincie_omschrijving,
  "Antwerpen" = "Antwerp",
  "Limburg" = "Limburg",
  "Oost-Vlaanderen" = "East Flanders",
  "Vlaams-Brabant" = "Flemish Brabant",
  "West-Vlaanderen" = "West Flanders",
  "Onbekend" = NA_character_
))
```

### municipality

```{r}
occurrence %<>% mutate(dwc_municipality = recode(gemeente_naam,
  "Onbekend" = NA_character_
))
```

### locationRemarks

```{r}
occurrence %<>% mutate(dwc_locationRemarks = recode(vha_categorie_omschrijving, 
  "CAT2 - Onbevaarbaar cat. 2" = "CAT2 - not navigable cat. 2",
  "CAT1 - Onbevaarbaar cat. 1" = "CAT1 - not navigable cat. 1",
  "CAT3 - Onbevaarbaar cat. 3" = "CAT3 - not navigable cat. 3",
  "CAT - Andere" = "CAT - other",
  "ONB - Onbekend" = "ONB - unknown",
  "BEV - Bevaarbaar" = "BEV - navigable"
))
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.5f",
  round(locatie_gps_breedte, digits = 5)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.5f",
  round(locatie_gps_lengte, digits = 5)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84") 
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```

### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### vernacularName

```{r}
occurrence %<>% mutate(dwc_vernacularName = vernacularName)
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

## Post-processing

```{r}
occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

```{r}
occurrence %<>% distinct(eventID, .keep_all = TRUE)
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-vmm-occurrences", "data", "processed", "occurrence.csv"), na = "")
```

# Checks

```{r}
test <- length(unique(occurrence$eventID))
test2 <- length(unique(occurrence$occurrenceID))

n_occur <- data.frame(table(occurrence$eventID))

```

