---
title: "Darwin Core Mapping"
subtitle: "MICA - Muskrat occurrences collected by RATO in East Flanders, Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}

#datafile V1
#input_interim <- read_csv(here::here("datasets", "mica-rato-occurrences", "data", "interim", "RATO_data_RAW_Refine.csv"))

#datafile update 1
input_interim <- read_excel(here::here("datasets", "mica-vmm-occurrences", "data", "raw","VangstenT14_47_33.xlsx"))

#input_interim <- read.csv2(here::here("datasets", "mica-rato-occurrences", "data", "raw", "RATOgegevensbackoffice2020V2.txt" #), sep = ";")

#deze werkt
#input_interim <- read_delim(file = here::here("datasets", "mica-rato-occurrences", "data", "raw", #"RATOgegevensbackoffice2020V2.txt"), ";", escape_double = FALSE, trim_ws = TRUE )

#input_interim <- read_delim(here::here("datasets", "mica-rato-occurrences", "data", "raw", #"RATOgegevensbackoffice2020V2.txt" ), ";")

#Vangsten_T11_16_45 <- read_excel("datasets/mica-vmm-occurrences/data/raw/VangstenT11_16_45.xlsx")

```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```
# Make speciesList


```{r}
speciesList <- input_interim %>%
  distinct(`Sporen Waarnemingen Naam`)  %>%
  pull() %>% # Create vector from dataframe
  parsenames()
head(speciesList, n = 30)
```

# filter for needed occurrences

```{r}
input_interim %<>%
        filter(`Sporen Waarnemingen Naam` != 'Vis') %<>%
        filter( `Sporen Waarnemingen Naam` != 'Andere') %<>%
        filter(`Sporen Waarnemingen Naam` != 'Eend' )%<>%
        filter(`Sporen Waarnemingen Naam` != 'Marterachtige' )
        
head (input_interim, n = 5 )
```





```{r}
input_interim %<>% mutate(vernacularName = `Sporen Waarnemingen Naam`)

input_interim %<>% mutate(vernacularName = recode(vernacularName
                   , "Muskusr. > 400gr"  = "Muskusrat"
                   , "Doodaars" = "Tachybaptus ruficollis"
                   
                                                  ))



input_interim %>% head(n = 5)
```

# Make speciesList2


```{r}
speciesList2 <- input_interim %>%
  distinct(`vernacularName`)  
  
head(speciesList2, n = 30)
```


# find scientificNames


```{r}
vernacular_names_df <- input_interim

get_vernacular_name <- function(vn) {
 names <-
   name_lookup(vn,
               datasetKey = "d7dddbf4-2cf0-4f39-9b2a-bb099caae36c",
               limit = 1)$data # this returns the most likely taxon
 if (nrow(names) > 0) {
   names$scientificName
 } else {
   NA_character_
 }
}

n_vernacular_names_df <-
 vernacular_names_df %>%
 group_by(vernacularName) %>%
 nest() %>%
 mutate(scientificName = map_chr(vernacularName,get_vernacular_name)) %>%
 ungroup() %>%
 select(-one_of("data")) %>%
 right_join(vernacular_names_df, by = "vernacularName")

view(n_vernacular_names_df)
```
# make speciesList

```{r}
speciesList3 <- n_vernacular_names_df %>%
  distinct(`scientificName`)  

input_interim <- n_vernacular_names_df    
head(speciesList3, n = 30)
view(input_interim)
```

# Occurrence core

look at interimfile once more

```{r}
head(input_interim, n = 15)
```


## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(Jaar, `Maand In Jaar`, `Maand Lange Omschrijving`, `Provincie NIS Code`))
```
## filter for muskrat

```{r eval=FALSE, include=FALSE}
occurrence %<>%
        filter(kind_nl == 'Bruine rat' | kind_nl == 'Muskusrat') 
        
head (occurrence, n = 5 )

```



# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "insert doi")
```

### type

```{r}
occurrence %<>% mutate(type = "Event")
```

### language

```{r}
occurrence %<>% mutate(language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(rightsHolder = "VMM") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(institutionCode = "INBO") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(datasetName = "MICA - Muskrat occurrences collected by RATO in East Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

```{r}
occurrence %<>% rename(occurrenceID = 'registratie ID') %>%
                mutate(occurrenceID = str_c("MICA:VMM:", occurrenceID))
  
```

### recordedBy

```{r}
occurrence %<>% mutate(recordedBy = 'VMM')
```

### individualCount

```{r}
occurrence %<>% rename(individualCount = 'vangst aantal') 
```

### organismQuantity

### organismQuentityType

### sex

### lifeStage

### behavior

### occurrenceRemarks

```{r eval=FALSE, include=FALSE}
occurrence %<>% rename(occurrenceRemarks = 'action_en')
```

## samplingProtocol

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(samplingEffort = materials_en)
```

```{r eval=FALSE, include=FALSE}
occurrence %<>% rename(samplingProtocol = 'rat_trap') 
```

### samplingEffort

```{r eval=FALSE, include=FALSE}
head(occurrence, n = 5)

occurrence %<>%
  # extract follow ups to new column
  mutate(samplingEffort = str_replace(samplingEffort, "Follow-up=*", "Times-visited=")) %>%
  mutate(samplingEffort = str_replace_all(samplingEffort, "=",":")) %>%
  mutate(samplingEffort = str_replace_all(samplingEffort, ","," |"))
```

### eventDate



```{r}


#occurrence %<>% ymd(date)

occurrence %<>% mutate(eventDate = (Maand))
# mutate(eventDate = as_date(ymd(date))) # , format = "%Y-%m-%d")) 

head(occurrence, n = 5)
```

# Location

```{r}
occurrence %<>%
  rename(decimalLongitude = X) %>%
  rename(decimalLatitude = Y) %>%
  mutate(geodeticDatum = "WGS84") %>%
  mutate(coordinateUncertaintyInMeters = "30") %>%
  mutate(verbatimCoordinateSystem = "Lambert coordinates") %>%
  mutate(verbatimSRS = "Belgian Datum 1972") %>%
  mutate(countryCode = "BE")  %>%            
  mutate(continent = "Europe")
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
  mutate(verbatimLongitude = round(verbatimLongitude)) %>%
  mutate(verbatimLatitude = round(verbatimLatitude)) %>%
  mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
  mutate(decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5)))
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>%
  mutate(municipality = str_remove(municipality,"\\(BE\\)"))
```

### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Lambert coordinates") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(identifiedBy = 'RATO')
```

### scientificName 

```{r}
occurrence %<>% mutate(scientificName = "Ondatra zibethicus")
```

### kingdom

```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(scientificNameAuthorship = "Linnaeus, 1766")
```

### taxonRank

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(occurrenceStatus = case_when(individualCount > 0 ~ "present",
                              individualCount == 0 ~ "absent"
                                 )
                                )
head(occurrence, n = 5)
```

## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
##occurrence %<>% select(-c(disposition, etiket_ref, collection, scientificName)) %<>% #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","occurrenceRemarks","samplingProtocol","samplingEffort"
               ,"eventDate","continent","countryCode","municipality"
               ,"verbatimLatitude","verbatimLongitude","verbatimCoordinateSystem","verbatimSRS"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","scientificNameAuthorship","taxonRank","nomenclaturalCode"
               )
occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-rato-occurrences", "data", "processed", " occurrence.csv"), na = "")
```
