---
title: "Darwin Core mapping"
subtitle: "For: Muskrat occurrences Flanders"
author:
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
```

# Read source data

Create a data frame `input_data` from the source data:

```{r}
input_data <- read_csv(file = here::here("data", "raw", "AgrDataGBIF.csv"), guess_max = 400000)
#input_data <- read_csv("../data/raw/AgrDataGBIF.csv")
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```

# Remove obsolete columns (Bijvangst)

```{r}
input_data %<>% select(-c(Bijvangst_soort_1, Bijvangst_soort_2, Bijvangst_soort_3, Bijvangst_aantal_1, Bijvangst_aantal_2, Bijvangst_aantal_3))
```

## Scientific names (not applicable)

Add columns:

```{r eval = FALSE}
input_data %<>%
  mutate(kingdom = "Animalia") %>%
  mutate(scientificName = "Ondatra zibethicus")
```

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the dataset:

```{r eval = FALSE}
parsed_names <- input_data %>%
  distinct(scientificName) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r eval = FALSE}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

Correct names and reparse:

```{r correct and reparse, eval = FALSE}
#input_data %<>% mutate(scientificName = recode(verbatimScientificName,
#  "Gibbulinella aff. dealbata" = "Gibbulinella dealbata",
#  "Hemicycla cf. gaudryi" = "Hemicycla gaudryi",
#  "Xerotricha aff. orbignii" = "Xerotricha orbignii",
#  "Napaeus cf. venegueraensis" = "Napaeus venegueraensis",
#  "Monilearia cf. praeposita" = "Monilearia praeposita",
#  "Gibbulinella aff. dewinteri" = "Gibbulinella dewinteri",
#  "Pomatias aff. laevigatus" = "Pomatias laevigatus",
#  "Physella cf. acuta" = "Physella acuta",
#  "Hemicycla psathyra cf. temperata" = "Hemicycla psathyra temperata"
#))

# Redo parsing
parsed_names <- input_data %>%
  distinct(scientific_name) %>%
  pull() %>%
  parsenames()

# Show names with nomenclatural issues again
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

## Taxon ranks (not applicable)

The nameparser function also provides information about the rank of the taxon (in `rankmarker`). Here we join this information with our checklist. Cleaning these ranks will done in the Taxon Core mapping:

```{r eval = FALSE}
input_data %<>% left_join(
  parsed_names %>%
  select(scientificname, rankmarker),
  by = c("scientificName" = "scientificname"))
```

```{r eval = FALSE}
# Adjust rankmarker
#input_data %<>% mutate(rankmarker = recode(verbatimScientificName,
#  "Monilearia spec." = "genus",
#  "Hemicycla spec." = "genus",
#  "Napaeus spec." = "genus",
#  .default = rankmarker
))
```

## Taxon IDs (not applicable)

To link taxa with information in the extension(s), each taxon needs a unique and relatively stable `taxonID`. Here we create one in the form of `dataset_shortname:taxon:hash`, where `hash` is unique code based on scientific name and kingdom (that will remain the same as long as scientific name and kingdom remain the same):

```{r eval = FALSE}
#vdigest <- Vectorize(digest) # Vectorize digest function to work with vectors
#input_data %<>% mutate(taxon_id = paste(
#  "muskrat-occurrences", # e.g. "alien-fishes-checklist"
#  "taxon",
#  vdigest(paste(scientificName, kingdom), algo = "md5"),
#  sep = ":"
#))
```

## Preview data

Show the number of taxa and distributions per kingdom and rank:

```{r}
#input_data %>%
#  group_by(kingdom, rankmarker) %>%
#  summarize(
#    `# taxa` = n_distinct(taxon_id),
#    `# distributions` = n()
#  ) %>%
#  adorn_totals("row")
```

Preview data:

```{r}
input_data %>% head()
```

# Darwin Core mapping

## Create occurrence core

Create a dataframe with occurrence data:

```{r}
occurrence <- input_data
```

## Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = recode(Inst,
  "Gem" = "municipality"
))
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "to complete") # e.g. "https://doi.org/10.15468/xvuzfh"
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Muskrat occurrences in Belgium, aggregated data") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

The following terms contain information about the taxon:

### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = paste0("INBO:MUSK:L:OCC:", X1))
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = dwc_rightsHolder)
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = Ratten)
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = case_when(
  !is.na(Datum) ~ as.character(Datum),
  !is.na(Maand) ~ as.character(paste(Jaar, sprintf("%02d", Maand), sep = "-")),
  !is.na(Jaar) ~ as.character(Jaar),
  TRUE ~ NA_character_
))
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = recode(Val_type,
  "Conibearklemmen" = "connibear clamp",
  "Klepfuiken" = "valve trap",
  "Lokaasklemmen"= "bait trap",
  "Ronde fuiken" = "round fikes",
  "Grondklemmen" = "ground trap"
))
```

### samplingEffort
 
```{r}
occurrence %<>% mutate(dwc_samplingEffort = case_when(
  !is.na(Mandagen) ~ paste("daysOfWork:", round(Mandagen, digits = 2)),
  !is.na(Val_aantal) ~ paste("numberOfTraps:", floor(Val_aantal)),
  TRUE ~ ""
))
```

### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = case_when(
  !is.na(VHAzone) ~ paste0("INBO:MUSK:L:LOC:VHA", VHAzone),
  TRUE ~ NA_character_
))
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE")
```

### stateProvince

```{r}
occurrence %<>% mutate(dwc_stateProvince = recode(Prov,
  "Antwerpen" = "Antwerp",
  "Vlaams Brabant" = "Flemish Brabant",
  "West-Vlaanderen"= "West Flanders",
  "Oost-Vlaanderen" = "East Flanders"
))
```

### Municipality

```{r}
occurrence %<>% mutate(dwc_municipality = recode(Gem,
  "Hélécine" = "Heylissem",
  "Crisnée" = "Gerstenhoven"
))
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = Lat)
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = Long)
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = case_when(
  endsWith(Coor_prox, "Yes") ~ "13000",
  endsWith(Coor_prox, "No") ~ "30",
  TRUE ~ NA_character_
))
```

### georeferenceRemarks

```{r}
occurrence %<>% mutate(dwc_georeferenceRemarks = case_when(
  endsWith(Coor_prox, "Yes") ~ "coordinates are centroid of municipality",
  endsWith(Coor_prox, "No") ~ "exact coordinates",
  TRUE ~ NA_character_
))
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = "Ondatra zibethicus")
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(dwc_scientificNameAuthorship = "(Linnaeus, 1766)")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

## Post-processing


```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")


```

remove obsolete columns

```{r}
occurrence %<>% select(-c(Coor_prox, Man_prox, Val_aantal, Grens, month, year, ownerInstitutionCode, Mandagen))
  
    
```




```{r}
```

remove NA

```{r}
#occurrence %<>% mutate(str_replace((samplingProtocol),"NA: NA", NA_character_))

occurrence %<>% mutate( samplingProtocol  =   str_replace((samplingProtocol),"NA: NA", NA_character_)) %>%
                mutate( samplingProtocol  =   str_replace((samplingProtocol),"NA", NA_character_)) %>%
                mutate(eventDate = str_replace((eventDate),"NA", NA_character_)) %>%
                mutate(samplingEffort = str_replace((samplingEffort),"NA", "")) %>%
                mutate(locationID = str_replace((locationID),"NA", NA_character_)) %>%
                mutate(municipality = str_replace((municipality),"NA", NA_character_)) 
                
                
                
  
                

```
```{r}
# occurrence %<>% mutate(eventDate = str_replace((eventDate),"NA", NA_character_))

```
```{r}
#occurrence %<>% mutate(samplingEffort = str_replace((samplingEffort),"NA", NA_character_))

```

```{r}
#occurrence %<>% mutate(locationID = str_replace((locationID),"NA", NA_character_))

```
```{r}
#occurrence %<>% mutate(municipality = str_replace((municipality),"NA", NA_character_))

```

Preview data:

```{r}
occurrence %>% head()
View(occurrence)
```

# Create presence only occurrence file
For now this is the one to be published

```{r}

presenceOnly <- filter(occurrence, individualCount > "0") 
#valType <- filter(occurrence , !is.na(samplingProtocol)) 
#andervalType <- occurrence %>% drop_na(Val_type)%>%
#val_Type <- filter(occurrence ,ValType != 'is.na') %>%
#  final <- occurrence[!(is.na(occurrence$Val_type)),]
#valtype <- filter(occurrence, !is.na(valtype),)
#finalData <- subset(occurrence,!(is.na(data["Val_type"])))
#vallekestype <- unique(occurrence$Val_aantal)
#View(valtype)
#View(vallekestype)

nAtest <- filter(occurrence ,locationID != 'is.na')
#View(nAtest)
```


Save to CSV:

```{r}
write_csv(occurrence, here("data", "processed", "occurrences_with_absence.csv"), na = "" )
write_csv(presenceOnly, here("data", "processed", "occurrence.csv"), na = "") 


#write_csv(occurrence,(file = here("data", "processed", "occurrence.csv"))
#write_csv(occurrence, here("data", "processed", "occurrence.csv"), append = FALSE, na = "")
#write.csv(occurrence, "../data/processed/occurrence.csv", append = TRUE, na = "",  row.names = FALSE, fileEncoding = "UTF-8")
```

